Versioned Queries on Native SPARQL Storage

### 2019-04-08 paper.tex
Small edits to intro and adding some more references

\begin{document}
\title{Versioned Queries in Native SPARQL Stores}
\author{Ignacio Cuevas \and Aidan Hogan}

\institute{IMFD Chile \& Department of Computer Science, University of Chile}

@@

\begin{abstract}
Although many prominent Linked Datasets are in a state of constant evolution, historical data are rarely published in a standard way. Though a number of works have recently addressed the issue of querying and managing archives of versioned RDF data, most propose dedicated solutions, which incurs a certain deployment cost. We rather propose solutions for representing archives of versioned RDF data using the standard SPARQL language, allowing existing implementations to be used. For querying, we consider two versioning primitives: viewing results from a particular version, or viewing changes between two consecutive versions. We identify and compare several models by which historical versions of an RDF graph can be stored as named graphs, and discuss how base queries can be rewritten in each model per the two aforementioned primitives. We then present a performance comparison of these models and baseline (non-)versioned queries with respect to 23 weekly versions of Wikidata and 232 user-defined queries, using Virtuoso as a reference SPARQL implementation.
\end{abstract}

\section{Introduction}
The Semantic Web~\footnote{\url{https://www.w3.org/2001/sw/wiki/Main_Page}} is an initiative of the World Wide Web Consortium, which envisions a ``Web of Data;'' that is, a web of \textit{Linked Data} where information has an standarized structure, making it easier for systems and machines to comunicate. In order to make such standarization possible, it becomes necessary to incorporate meta-data that describes the content, meaning and relationship of data in all databases.

The Resource Description Framework~\cite{key:rdfprimer11} (RDF for short) is a standard model for data interchange on the Web. Its design is based on ``triples,'' which represent connections on a graph, where every node is a ``resource'' with ``predicates'' as the edges connecting them. Both resources and predicates are formally defined, and a collection of rules allow for inferences about the relationships in the data.
Alongside RDF, the ontological languages RDFS~\cite{key:rdfschema11} and OWL~\cite{key:owl2primer} provide basic relation and type definitions, and SPARQL~\cite{key:sparql11} serves as RDF's query engine (similar to SQL on relational databases).

@@

The prior three works share their focus on formal definitions over implementations. They also do not consider eficiency or scalability analysis and employ SPARQL extensions and/or specialized indices. In light of these facts, said works will be considered only to compare results for the most part.

>>>>

\begin{document}
\title{Versioned Queries on Native SPARQL Storage}
%\author{Ignacio Cuevas \and Aidan Hogan}

%\institute{IMFD Chile \& Department of Computer Science, University of Chile}

@@

\begin{abstract}
Although many prominent Linked Datasets are in a state of constant evolution, historical data are rarely published in a standard way. Though a number of works have recently addressed the issue of querying and managing archives of versioned RDF data, most propose dedicated solutions, which incurs a high deployment cost. We rather propose solutions for representing archives of versioned RDF data using the standard SPARQL language, allowing existing implementations to be used. For querying, we consider two versioning primitives: viewing results from a particular version, or viewing changes between two consecutive versions. We identify and compare several models by which historical versions of an RDF graph can be stored as named graphs, and discuss how base queries can be rewritten in each model per the two aforementioned primitives. We then present a performance comparison of these models and baseline (non-)versioned queries with respect to 23 weekly versions of Wikidata and 232 user-defined queries, using Virtuoso as a reference SPARQL implementation.
\end{abstract}

\section{Introduction}

A key aspect of the Web is its dynamic nature, where documents are frequently updated, deleted and added. Likewise when we speak of the Semantic Web, it is important to consider that sources may be dynamic and RDF datasets are subject to change. The traditional Web and the Semantic Web thus share similar strengths and limitations regarding the dynamics of information. In terms of strengths, the flexible nature of both webs means that documents can be updated with little restriction and with little need for centralised coordination. On the other hand, in terms of limitations, neither web has built in support for preservation, nor for versioning; for example, on the traditional Web, the HTTP protocol does not permit requesting a past version of a webpage. For this reason, a number of specialised ``Web archives'' have emerged that attempt to capture and track different intermittent versions of documents on the Web, the most prominent of which is the Internet Archive~\cite{JaffeK09}. While similar techniques can be applied to Semantic Web documents containing RDF data -- simply archiving and thus preserving the syntax of the document itself -- the structured nature of such content means that an RDF archive can potentially do a lot more.

A number of works have looked at archiving RDF content from the Web. One such example is the Dynamic Linked Data Observatory (DyLDO)~\cite{KaferAUOH13}, which has been tracking and archiving changes in a sample of around 100,000 RDF documents each week since 2013. Other works rather focus on aspects relating to detecting change~\cite{TummarelloMBE07,ZeginisTC11,PapavasileiouFFKC13,KaferAUOH13,DividinoKG14,RoussakisCSFS15,NishiokaS18}, notifying of changes~\cite{TummarelloMBE07,PopitschH11,PassantM10,TrampFEA10}, versioning data~\cite{VolkelG06,GraubeHU14,KhuranaD16}, and so forth.

% RDF archives~\cite{FernandezUPK19}

@@

The prior three works share their focus on formal definitions over implementations. They also do not consider efficiency or scalability analysis and employ SPARQL extensions and/or specialized indices. In light of these facts, said works will be considered only to compare results for the most part.
